%!TEX root = thesis.tex
\chapter{\label{cha:Preface}Preface}

This report is the result of my Master's thesis project performed at the Software Engineering Research Group (SERG) at Delft University of Technology. With this project I got the opportunity to research an aspect of software engineering I find most interesting: combining historical data to assist in every day software engineering processes, such as bug triaging.

Crafting this thesis took quite a while. Not only did I explore several other possibilities for a thesis project (that ended up not working out), also working on a thesis project \emph{and} running a successful company was really challenging. It was always a struggle to choose between business and thesis project, resulting quite often in a lack of focus.

But finally, after some hard work, focus, and determination, before you lies the result of my work. First of all, I would like to thank Ali Mesbah for supervising me during the first part of my literature research. After Ali left to pursue a career at the University of British Columbia, Martin Pinzger became my daily supervisor. Martin, first of all thank you for all your patience with me, it has been quite a ride. I really appreciated the interesting discussions we had and the guidance, support and feedback you offered me. Finally, thank you for keeping me focused towards my goal: finishing this thesis. 

To all others that helped me during my thesis project, in whatever manner, thank you for all your support and patience. 

\vskip1cm
\begin{flushright}
\theauthor\\
Delft, The Netherlands \\
August 17, 2012\\
\end{flushright}

